---
title: "Make Lapses" 
author: "Kendra Wyant"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---   

### Notes

This script uses `study_dates.csv` from the gps study and the cleaned `lapses.csv` RISK file. It filters down to lapses that are longer than one hour and outputs a tibble containing `subid`, `lapse_start`, and `lapse_end`.    

This file will be used to filter down gps data to location during lapses and used for risk terrain modeling. 

### Outstanding Descisions

Most lapses are around 4 hours or less. Consider excluding unusually long lapses (e.g., 24+ hours).


### Setup

```{r}
#| message: false

options(conflicts.policy = "depends.ok")

library(tidyverse)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")

path_processed <- format_path("studydata/risk/data_processed/shared")
path_gps <- format_path("studydata/risk/data_processed/gps")
path_out <- format_path("studydata/risk/data_processed/terrain")

```


### Read in lapses

```{r}
lapses <- read_csv(here::here(path_processed, "lapses.csv"),
                show_col_types = FALSE) |> 
  glimpse()
```

### Read in Study Dates

```{r}
study_dates <- read_csv(here::here(path_gps, "study_dates.csv"),
                show_col_types = FALSE) |> 
  glimpse()
```

### Filter Lapses

Filter to subids in `study_dates.csv`    
- only 45 lapses lost by this filtering
```{r}
lapses <- lapses |> 
  filter(subid %in% study_dates$subid) |> 
  glimpse()
```

Filter to lapses at least 1 hour in duration  
- 257 lapses lost in this filtering, but we still have a good amount
```{r}
lapses <- lapses |> 
  filter(duration >= 1) |> 
  glimpse()
```


### Light EDA

774 lapses from 71 unique subids
```{r}
lapses |> nrow()

lapses |> 
  group_by(subid) |> 
  count() |> 
  arrange(desc(n))
```

Some extra long lapses - we might need to exclude?
```{r}
lapses |> 
  ggplot(aes(duration)) +
  geom_histogram() +
  labs(title = "Lapse Duration Distribution",
       x = "Duration (hours)",
       y = "Count")
```

12 lapses over 24 hours. Could consider excluding these. Will retain for now
```{r}
lapses |> 
  filter(duration > 24) |> 
  arrange(desc(duration)) |> 
  select(subid, duration)
```

It looks like 75% of lapses are 4 hours or less in duration
```{r}
lapses |> 
  skimr::skim(duration)
```


### Save lapse file
```{r}
lapses |> 
  select(subid, lapse_start, lapse_end) |> 
  write_csv(here::here(path_out, "lapses.csv")) |> 
  glimpse()
```
